Learning to Pivot with Adversarial Networks

### 2017-05-15 tex/nips/nips_2017.tex
Use NIPS template

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ICML 2016 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use the following line _only_ if you're still using LaTeX 2.09.
%\documentstyle[icml2016,epsf,natbib]{article}
% If you rely on Latex2e packages, like most moden people use this:
\documentclass{article}

\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}

% use Times
\usepackage{times}
% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

@@

% For citations
\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}

% As of 2011, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2016} with
% \usepackage[nohyperref]{icml2016} above.
\usepackage{hyperref}

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''
\usepackage{icml2016}

% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Proceedings of the...''
%\usepackage[accepted]{icml2016}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Learning to Pivot with Adversarial Networks}

\begin{document}

\twocolumn[
\icmltitle{Learning to Pivot with Adversarial Networks}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2016
% package.
\icmlauthor{Gilles Louppe}{g.louppe@nyu.edu}
\icmladdress{New York University}
\icmlauthor{Michael Kagan}{makagan@slac.stanford.edu}
\icmladdress{SLAC National Accelerator Laboratory}
\icmlauthor{Kyle Cranmer}{kyle.cranmer@nyu.edu}
\icmladdress{New York University}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
%\icmlkeywords{boring formatting information, machine learning, ICML}

\vskip 0.3in
]

\begin{abstract}

Many inference problems involve data generation processes that are not uniquely
specified or are uncertain in some way. In a scientific context, the presence of
several plausible data generation processes is often associated to the presence
of systematic uncertainties. Robust inference is possible if it is based on a
pivot -- a quantity whose distribution does not depend on the unknown value of the
(categorical or continuous) nuisance parameters that parametrizes this family of
generation processes. In this work,  we introduce a flexible training procedure
based on adversarial networks for enforcing the pivotal property on a predictive model, i.e., its independence
with respect to nuisance parameters. We derive theoretical results showing that the proposed algorithm tends
towards a minimax solution corresponding to a predictive model that is both
optimal and independent of the nuisance parameters (if that model exists) or
for which one can tune a trade-off between power and robustness. Finally, we
demonstrate the effectiveness of this approach with a toy example and an example
from particle physics.

\end{abstract}

@@

    \begin{tikzpicture}[shorten >= 1pt, ->, node distance=\layersep,scale=.85]
    \tikzstyle{neuron} = [circle, minimum size=0.25cm, draw=black!20, line width=0.3mm, fill=white]

@@

    \begin{flushleft}
        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;\\
        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$;\\
        {\it Hyper-parameters:} Number $T$ of training iterations,
                                Number $K$ of gradient steps to update $r$,
                                Size $M$ of the minibatches.
    \end{flushleft}

@@

\begin{figure*}
    \begin{center}
        \includegraphics[width=0.245\textwidth]{../figures/f-plain.pdf}
        \includegraphics[width=0.245\textwidth]{../figures/surface-plain.pdf}
        \includegraphics[width=0.245\textwidth]{../figures/f-adversary.pdf}
        \includegraphics[width=0.245\textwidth]{../figures/surface-adversary.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{Toy example.
    (Left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$
       when $f$ is built without adversarial training. The resulting densities
       are clearly dependent on the continuous parameter $Z$, indicating that $f$ is not pivotal.
    (Middle left) The corresponding decision surface, highlighting
       the fact that samples are easier to classify for values of $Z$ above $\sigma$,
       hence partially explaining the dependency.
    (Middle right) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$ when $f$ is
       built with adversarial training, as outlined in Sec.~\ref{sec:method}.
       The resulting densities are now almost identical to each other, indicating only a
       small dependency on $Z$.
    (Right) The corresponding decision surface, illustrating how adversarial
       training bends the decision function vertically to erase the dependency on $Z$.
    }
    \label{fig:toy}
\end{figure*}

@@

\begin{figure}
    \includegraphics[width=0.48\textwidth]{../figures/training.pdf}
    \vspace{-1cm}
    \caption{Toy example. Training curves for ${\cal L}_f(\theta_f)$, ${\cal L}_r(\theta_f, \theta_r)$
             and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
             Initialized with a pre-trained classifier $f$, adversarial training was performed for $200$ iterations, mini-batches of size $M=128$, $K=500$ and $\lambda=50$.}
    \label{fig:toy-training}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=0.48\textwidth]{../figures/ams.pdf}
    \caption{Physics example. Approximate median significance as a function of the decision threshold
             on the output of $f$.  As shown at $\lambda=10$, trading
             classification accuracy for independence with respect to pileup
             results in a net benefit in terms of statistical significance.}
    \label{fig:physics-ams}
\end{figure}

@@

% \section*{Acknowledgements}
%
% We would like to thank the authors of \citep{baldi2016jet} for sharing the data used in their
% studies.
% KC and GL are both supported through NSF ACI-1450310, additionally KC is supported
% through PHY-1505463 and PHY-1205376.  MK is supported by the US
% Department of Energy (DOE) under grant DE-AC02-76SF00515 and by the
% SLAC Panofsky Fellowship.

% In the unusual situation where you want a paper to appear in the
% references without citing it in the main text, use \nocite

\bibliography{bibliography}
\bibliographystyle{icml2016}

@@

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was
% created by Lise Getoor and Tobias Scheffer, it was slightly modified
% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz,
% slightly modified from the 2009 version by Kiri Wagstaff and
% Sam Roweis's 2008 version, which is slightly modified from
% Prasad Tadepalli's 2007 version which is a lightly
% changed version of the previous year's version by Andrew Moore,
% which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

>>>>

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

%\usepackage{nips_2017}

% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}

@@

%\newcommand{\theHalgorithm}{\arabic{algorithm}}

\title{Learning to Pivot with Adversarial Networks}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  Gilles Louppe\\
  New York Univeristy\\
  \texttt{g.louppe@nyu.edu}\\
  \And
  Michael Kagan\\
  SLAC National Accelerator Laboratory\\
  \texttt{makagan@slac.stanford.edu}\\
  \And
  Kyle Cranmer\\
  New York Univeristy\\
  \texttt{kyle.cranmer@nyu.edu}
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
    Many inference problems involve data generation processes that are not uniquely
    specified or are uncertain in some way. In a scientific context, the presence of
    several plausible data generation processes is often associated to the presence
    of systematic uncertainties. Robust inference is possible if it is based on a
    pivot -- a quantity whose distribution does not depend on the unknown value of the
    (categorical or continuous) nuisance parameters that parametrizes this family of
    generation processes. In this work,  we introduce a flexible training procedure
    based on adversarial networks for enforcing the pivotal property on a predictive model, i.e., its independence
    with respect to nuisance parameters. We derive theoretical results showing that the proposed algorithm tends
    towards a minimax solution corresponding to a predictive model that is both
    optimal and independent of the nuisance parameters (if that model exists) or
    for which one can tune a trade-off between power and robustness. Finally, we
    demonstrate the effectiveness of this approach with a toy example and an example
    from particle physics.
\end{abstract}

@@

    \begin{tikzpicture}[shorten >= 1pt, ->, node distance=\layersep,scale=.65]
    \tikzstyle{neuron} = [circle, minimum size=0.25cm, draw=black!20, line width=0.3mm, fill=white]

@@

    \begin{flushleft}
        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;
        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$.
        % {\it Hyper-parameters:} Number $T$ of iterations,
        %                         Number $K$ of gradient steps to update $r$,
        %                         Size $M$ of the minibatches.
    \end{flushleft}

@@

@@

\begin{figure*}
    \begin{center}
        \includegraphics[width=0.245\textwidth]{../figures/f-plain.pdf}
        \includegraphics[width=0.245\textwidth]{../figures/surface-plain.pdf}
        \includegraphics[width=0.245\textwidth]{../figures/f-adversary.pdf}
        \includegraphics[width=0.245\textwidth]{../figures/surface-adversary.pdf}
    \end{center}
    \caption{Toy example.
    (Left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$
       when $f$ is built without adversarial training. The resulting densities
       are clearly dependent on the continuous parameter $Z$, indicating that $f$ is not pivotal.
    (Middle left) The corresponding decision surface, highlighting
       the fact that samples are easier to classify for values of $Z$ above $\sigma$,
       hence partially explaining the dependency.
    (Middle right) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$ when $f$ is
       built with adversarial training, as outlined in Sec.~\ref{sec:method}.
       The resulting densities are now almost identical to each other, indicating only a
       small dependency on $Z$.
    (Right) The corresponding decision surface, illustrating how adversarial
       training bends the decision function vertically to erase the dependency on $Z$.
    }
    \label{fig:toy}
\end{figure*}

\begin{figure}
\centering
\begin{minipage}{.46\linewidth}
    \centering
        \includegraphics[width=0.95\textwidth]{../figures/training.pdf}
        \caption{Toy example. Training curves for ${\cal L}_f(\theta_f)$, ${\cal L}_r(\theta_f, \theta_r)$
                 and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
                 Initialized with a pre-trained classifier $f$, adversarial training was performed for $200$ iterations, mini-batches of size $M=128$, $K=500$ and $\lambda=50$.}
        \label{fig:toy-training}
\end{minipage}
\hspace{.05\linewidth}
\begin{minipage}{.46\linewidth}
    \centering
        \includegraphics[width=0.95\textwidth]{../figures/ams.pdf}
        \caption{Physics example. Approximate median significance as a function of the decision threshold
                 on the output of $f$.  As shown at $\lambda=10$, trading
                 classification accuracy for independence with respect to pileup
                 results in a net benefit in terms of statistical significance.}
        \label{fig:physics-ams}
\end{minipage}
\end{figure}

@@

@@

\bibliography{bibliography}
\bibliographystyle{apalike}

@@
